### 2022-09-15 main/chapter_searches.tex
minor stats tweaks

Confidence intervals are confusing.
It is easy to be tricked into thinking that the confidence interval for given
data `probably' contains the unknown parameter, but that gets it backwards.
For a known true parameter $x$, the confidence intervals in random repeated
experiments will contain that $x$ with probability $1 - \alpha$;
the interval makes a subtle statement about unknown alternative data for
known true parameters.
Given confidence intervals can even exclude all plausible parameters~\cite{
pratt1961testing,
Jaynes1976intervals
}.
Pratt eloquently explains the correct interpretation of a confidence interval
by example:
\begin{displayquote}
\small
``%
A method yielding true statements with probability $.95$, when applied to your
experiment, yields the statement that your treatment effect is between $17$ and
$29$, but no conclusion is possible about how probable it is that your
treatment effect is between $17$ an $29$.%
''~\cite{pratt1961testing}.
\end{displayquote}
Neyman and Pearson never claimed otherwise; they did precise mathematics, and
confusion about their intervals arises externally~\cite{jaynes2003probability}.

Confidence intervals are functions of the data, so can inform vague inferences
if used as data for likelihood functions, just as we did to interpret Fisher's
\pvalue\ ratio as a likelihood ratio ---
an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
likelihood function
\begin{equation}
L(x) =
\left\{
\begin{matrix}
1 - \alpha & \textrm{if}~x \in x_1\textrm{--}x_2, \\
\alpha & \textrm{otherwise.} \\
\end{matrix}
\right.
\end{equation}
And again, this reduction cannot add to the data.
It can only detract.
The recommendation by Fisher to directly communicate the likelihood function
remains simpler and clearer.
In response to the idea that no conclusion can be drawn about the parameter
iself, Fisher also deserves the last word:
\begin{displayquote}
\small
``%
There is something horrifying in the ideological movement represented by the
doctrine that reasoning, properly speaking, cannot be applied to empirical data
to lead to inferences valid in the real world.%
''~\cite{fisher1956statistical}
\end{displayquote}

@@

Such misrepresentations of our results reinforces the need for clearer
replacements, which $\log$ likelihood ratios can provide.
And standard practice at the LHC contains the beginnings of such
replacements under the name $\cls$~\cite{
read2000modified,
Read2002cls,
junk1999confidence,
read1997optimal,
bock1998lower,
etde1998prospects,
lep2000searches,
lep2003search,
Murray2010heretic,
CMS-NOTE-2011-005,
pdg2022ynf
},
which is used for setting upper limits on disfavoured signal yields $s$
above background yields $b$;
\begin{equation}
\label{eqn:searches_cls}
\cls
=
\frac{P(s + b)}{P(b)\hphantom{s +\;}}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
}
,
\end{equation}
in which the individual \pvalues\ are named
$\clb = P(b)$ and
$\clsb = P(s + b)$.%
\footnote{%
Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
not confidence levels ---
a confidence level $\gamma$ is a decision boundary used, for example,
to assign limits by solving for $x$ such that
$\gamma = 1 - P(x)$~\cite{pdg1996}.
In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
by~\cite{pdg1984}) do make the distinction~\cite{
eadie1971statistical,
frodesen1979probability
}.%
}%
\footnote{%
Some literature defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
}, but such a $p_b$ might not be a \pvalue\ if it excludes equality with the
observed test statistic.
}%
As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
on obfuscated data, so (with some caution) $\cls$ can be interpreted as a
likelihood ratio.%
\footnote{%
Likelihood ratios are useful if both their numerator and denominator use the
same data, so $\clsb$ and $\clb$ should use the same test statistic.
Identical test statistics are used for all results in presented in this thesis,
but differing test statistics have been allowed elsewhere~\cite{bock1998lower}.
}%
At a given confidence level $\gamma$ (typically $95\%$), limits are set to
values of $s$ for which $\cls = 1 - \gamma$.

>>>>

Confidence intervals are confusing.
It is easy to be tricked into thinking that the confidence interval for given
data `probably' contains the unknown parameter, but that gets it backwards.
For a known true parameter $x$, the confidence intervals in random repeated
experiments will contain that $x$ with probability $\gamma$;
the interval makes a subtle statement about unknown alternative data for
known true parameters.
Given confidence intervals can even exclude all plausible parameters~\cite{
pratt1961testing,
Jaynes1976intervals
}.
Pratt eloquently explains the correct interpretation of a confidence interval
by example:
\begin{displayquote}
\small
``%
A method yielding true statements with probability $.95$, when applied to your
experiment, yields the statement that your treatment effect is between $17$ and
$29$, but no conclusion is possible about how probable it is that your
treatment effect is between $17$ an $29$.%
''~\cite{pratt1961testing}.
\end{displayquote}
Neyman and Pearson never claimed otherwise; they did precise mathematics, and
confusion about their intervals arises externally~\cite{jaynes2003probability}.

Confidence intervals are functions of the data, so can inform vague inferences
if used as data for likelihood functions, just as we did to interpret Fisher's
\pvalue\ ratio as a likelihood ratio ---
an accurately constructed confidence interval $x_1\textrm{--}x_2$ has a
likelihood function
\begin{equation}
L(x) =
\left\{
\begin{matrix}
\gamma & \textrm{if}~x \in x_1\textrm{--}x_2, \\
1 - \gamma & \textrm{otherwise.} \\
\end{matrix}
\right.
\end{equation}
And again, this reduction cannot add to the data.
It can only detract.
The recommendation by Fisher to directly communicate the likelihood function
remains simpler and clearer.
In response to the idea that no conclusion can be drawn about the parameter
iself, Fisher also deserves the last word:
\begin{displayquote}
\small
``%
There is something horrifying in the ideological movement represented by the
doctrine that reasoning, properly speaking, cannot be applied to empirical data
to lead to inferences valid in the real world.%
''~\cite{fisher1956statistical}
\end{displayquote}

@@

Such misrepresentations of our results reinforces the need for clearer
replacements, which $\log$ likelihood ratios can provide.
And standard practice at the LHC contains the beginnings of such
replacements under the name $\cls$~\cite{
read2000modified,
Read2002cls,
junk1999confidence,
read1997optimal,
bock1998lower,
etde1998prospects,
lep2000searches,
lep2003search,
Murray2010heretic,
CMS-NOTE-2011-005,
pdg2022ynf
},
which is used for setting upper limits on disfavoured signal yields $s$
above background yields $b$;
\begin{equation}
\label{eqn:searches_cls}
\cls
=
\frac{P(s + b)}{P(b)\hphantom{s +\;}}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}\hphantom{s +\;}
}
,
\end{equation}
in which the individual \pvalues\ are named
$\clb = P(b)$ and
$\clsb = P(s + b)$.%
\footnote{%
Despite the $\mathrm{CL}$ labels, these ($\cls$, $\clsb$, and $\clb$) are
not confidence levels ---
a confidence level $\gamma$ is a decision boundary used, for example,
to assign limits by solving for $x$ such that
$\gamma = 1 - P(x)$~\cite{pdg1996}.
In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
by~\cite{pdg1984}) do make the distinction~\cite{
eadie1971statistical,
frodesen1979probability
}.%
}%
\footnote{%
Some literature defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
}, but such a $p_b$ might not be a \pvalue\ when it excludes equality with the
observed test statistic~\cite{CMS-NOTE-2011-005}.
}%
As discussed in Section~\ref{sec:searches_fisher}, \pvalues\ are likelihoods
on obfuscated data, so (with some caution) $\cls$ can be interpreted as a
likelihood ratio.%
\footnote{%
Likelihood ratios are useful if both their numerator and denominator use the
same data, so $\clsb$ and $\clb$ should use the same test statistic.
Identical test statistics are used for all results in presented in this thesis,
but differing test statistics have been allowed elsewhere~\cite{bock1998lower}.
}%
At a given confidence level $\gamma$ (typically $95\%$), limits are set to
values of $s$ for which $\cls = 1 - \gamma$.
